Typed Functional Genetic Programming

### 2013-06-22 tex/thesis.tex
[tex] GP pseudocode...

\usepackage[utf8]{inputenc}
\usepackage{qtree}
\usepackage{xspace}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amsthm}
\usepackage{graphicx}
\usepackage{hyperref}

@@

@@

	\textit{Genetic programming} (GP) is a technique inspired by biological evolution
	that for a given problem tries to find computer programs able to solve that problem.

	A problem to be solved is given to GP in a form of \textit{fitness function}. 
	Fitness function is a function which takes computer program as its input and 
	returns numerical value called \textit{fitness} as output. 
	The bigger fitness of a computer program is, the better solution of a problem it is.

@@

	\Lets describe GP algorithm briefly.

	At the beginning initial population is generated randomly.

@@

	This step consists of two sub steps\footnote{TODO : Technically it is done in a little  
	bit different fashion which is equivalent.}:
	\begin{itemize} 
		\item Selection of \textit{parents} for individuals of the next population based on the   		          fitness.
		      The bigger fitness of an individual of the current population is, 
		      the better chance of success being selected as parent it has.  
		\item Application of genetic operators (such as \textit{crossover}, 
		      \textit{reproduction} and \textit{mutation}) 
			  on parent individuals producing new individuals of the next population.  
	\end{itemize}	  
	This transformation is repeatedly applied for a predefined number of steps (which is called 
	number of \textit{generations}) or until some predefined criterion is met.

	...
	\\\\
	\textbf{TODO}
	\begin{itemize} 
		\item TALK ABOUT GP is part of EA etc. and maybe define the GP by defining EA and then
		      specifying the differences or something like that...    
		\item History, citations, etc ....
	\end{itemize}

@@

@@

\bibitem{barendregt10}
  Henk Barendregt, Wil Dekkers, Richard Statman,
  \emph{Lambda Calculus With Types}.
  Cambridge University Press,
  2010. \\
  \url{http://www.cs.ru.nl/~henk/book.pdf}

>>>>

\usepackage[utf8]{inputenc}
\usepackage{qtree}

\usepackage[vlined]{algorithm2e}
%% \usepackage{algpseudocode}
\usepackage{framed}

\usepackage{xspace}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amsthm}
\usepackage{graphicx}
\usepackage{hyperref}

@@

\newcommand{\Pseudokod}[2]{
	\begin{framed}
	\begin{algorithm}[H]
		\DontPrintSemicolon
		\SetKwProg{Fn}{Algorithm}{}{}
		\Fn{#1}{#2}
	\end{algorithm}
	\end{framed}
}

@@

\textit{Genetic programming} (GP) is a technique inspired by biological evolution
that for a given problem tries to find computer programs able to solve that problem. 
GP was developed by John Koza \cite{koza92} in 1992.

A problem to be solved is given to GP in a form of \textit{fitness function}. 
Fitness function is a function which takes computer program as its input and 
returns numerical value called \textit{fitness} as output. 
The bigger fitness of a computer program is, the better solution of a problem.

@@

Individuals are computer program \textit{expressions} kept as \textit{syntactic trees}. 
Basically those trees are rooted trees with a function symbol in each internal node 
and with constant symbol or variable symbol in each leaf node. 
Number of child nodes for each internal node corresponds to the number of arguments of a function whose
symbol is in that node.

Another crucial input besides fitness function is a collection of \textit{building blocks}.
It is collection of symbols (accompanied with an information about number of arguments).
Those symbols are used to construct trees representing individuals.  
\\\\
\Lets describe GP algorithm briefly:

At the beginning initial population is generated from building blocks randomly.

@@

This step consists of two sub steps\footnote{TODO : Technically it is done in a little  
bit different fashion which is equivalent.}:
\begin{itemize} 
	\item Selection of \textit{parents} for individuals of the next population based on the fitness.
	      The bigger fitness of an individual of the current population is, 
	      the better chance of success being selected as parent it has.  
	\item Application of genetic operators (such as \textit{crossover}, 
	      \textit{reproduction} and \textit{mutation}) 
		  on parent individuals producing new individuals of the next population.  
\end{itemize}	  
This transformation is repeatedly applied for a predefined number of steps (which is called 
number of \textit{generations}) or until some predefined criterion is met.	
\\\\
\Lets now look on GP at more detail. Here is pseudocode for GP algorithm:

\Pseudokod{GP(fitnessFun,buildingBlocks,numGens,popSize,probabs)}{
	$gen \leftarrow$ 0 \;
	$pop \leftarrow generateInitialPopuletion( buildingBlocks ) $ \;
	($popWithFitnesses,terminate) \leftarrow$ evaluate($fitnessFun$, $pop$) \;

	\While{ $gen < numGens$ $\wedge$ $\neg terminate$  }{

	$newPop \leftarrow$ empty population \;
	$i \leftarrow 0$ \;

	\While{ $i < popSize$ }{
		$op \leftarrow probabilisticallySelectOperation(probabs)$ \;		
		\Switch{op}{
			\Case{Crossover}{
				$indiv1 \leftarrow$ selection( $popWithFitnesses$ ) \;
				$indiv2 \leftarrow$ selection( $popWithFitnesses$ ) \;
				($child1$,$child2$) = crossover( $indiv1$ , $indiv2$ ) \;
				$newPop$.insert( $child1$ ) \;
				$newPop$.insert( $child2$ ) \;
				$i \leftarrow i + 2$ \;
			}
			\Case{Reproduction}{
				$indiv \leftarrow$ selection( $popWithFitnesses$ ) \;
				$newPop$.insert( $indiv$ ) \;
				$i \leftarrow i + 1$ \;
			}
			\Case{Mutation}{
				$indiv \leftarrow$ selection( $popWithFitnesses$ ) \;
				$mutant \leftarrow$ mutate( $indiv$ ) \;
				$newPop$.insert( $mutant$ ) \;
				$i \leftarrow i + 1$ \;			
			}		
		}
	}

	$pop \leftarrow newPop$  \;
	$gen \leftarrow gen + 1$ \;
	\;
	($popWithFitnesses,terminate) \leftarrow$ evaluate($fitnessFun$, $pop$) \; 
 }
 \Return pop \;
}

In order to clarify the code let us describe in greater detail inputs and
contained procedures
generateInitialPopuletion(),
evaluate(),
probabilisticallySelectOperation(),
selection(),
crossover() 
and mutation().

@@

    ...
	\\\\
	\textbf{TODO}
	\begin{itemize} 
		\item TALK ABOUT GP is part of EA etc. and maybe define the GP by defining EA and then
		      specifying the differences or something like that...    
		\item History, citations, etc ....
		\item ten algoritmus v kozovi dělá "Designate Result" já tam vracim poslední populaci
	\end{itemize}

@@

\bibitem{koza92}

  John R. Koza,
  \emph{Genetic Programming: On the Programming of Computers by Means of Natural Selection}.
  MIT Press, Cambridge, MA,
  1992.

\bibitem{barendregt10}

  Henk Barendregt, Wil Dekkers, Richard Statman,
  \emph{Lambda Calculus With Types}.
  Cambridge University Press,
  2010. \\
  \url{http://www.cs.ru.nl/~henk/book.pdf}
